# Reproducible Research: Project 1

```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
library(ggplot2)
library(scales)
library(Hmisc)
```

## Loading and preprocessing the data
##### 1. Load the data
```{r, results='markup', warning=TRUE, message=TRUE}

Activity_Data <- read.csv('activity.csv')
```

-----

## What is mean total number of steps taken per day?
```{r}
Steps_By_Day <- tapply(Activity_Data$steps, Activity_Data$date, sum, na.rm=TRUE)
```

##### 1. Make a histogram of the total number of steps taken each day
```{r}
qplot(Steps_By_Day, xlab='Total Steps', ylab='Frequency', binwidth=500)
```

##### 2. Calculate and report the mean and median total number of steps taken per day
```{r}
Steps_By_Day_Mean <- mean(Steps_By_Day)
Steps_By_DayMedian <- median(Steps_By_Day)
```
* Mean: `r Steps_By_Day_Mean`
* Median:  `r Steps_By_DayMedian`

-----

## What is the average daily activity pattern?
```{r}
Average_Steps_Per_Time_Block <- aggregate(x=list(meanSteps=Activity_Data$steps), by=list(interval=Activity_Data$interval), FUN=mean, na.rm=TRUE)
```

##### 1. Make a time series plot
```{r}
ggplot(data=Average_Steps_Per_Time_Block, aes(x=interval, y=meanSteps)) +
    geom_line() +
    xlab("Interval") + 
    ylab("Number Of Steps") 
```

##### 2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
Most_Steps <- which.max(Average_Steps_Per_Time_Block$meanSteps)
Time_Most_Steps <-  gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", Average_Steps_Per_Time_Block[Most_Steps,'interval'])
```

* Most Steps at: `r Time_Most_Steps`

----

## Imputing missing values
##### 1. Calculate and report the total number of missing values in the dataset 
```{r}
Number_Of_Missing_Values <- length(which(is.na(Activity_Data$steps)))
```

* Number of missing values: `r Number_Of_Missing_Values`

##### 2. Devise a strategy for filling in all of the missing values in the dataset.
##### 3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
Activity_Data_Altered <- Activity_Data
Activity_Data_Altered$steps <- impute(Activity_Data$steps, fun=mean)
```


##### 4. Make a histogram of the total number of steps taken each day 
```{r}
Steps_By_Day_Altered <- tapply(Activity_Data_Altered$steps, Activity_Data_Altered$date, sum)
qplot(Steps_By_Day_Altered, xlab='Total Steps', ylab='Frequency', binwidth=500)
```

##### ... and Calculate and report the mean and median total number of steps taken per day. 
```{r}
Steps_By_Day_Mean_Altered <- mean(Steps_By_Day_Altered)
Steps_By_Day_Median_Altered <- median(Steps_By_Day_Altered)
```
* Mean (Altered): `r Steps_By_Day_Mean_Altered`
* Median (Altered):  `r Steps_By_Day_Median_Altered`

----

## Are there differences in activity patterns between weekdays and weekends?
##### 1. Create a new factor variable in the dataset with two levels – “Week Day” and “Week End” indicating whether a given date is a weekday or weekend.

```{r}
Activity_Data_Altered$dateType <-  ifelse(as.POSIXlt(Activity_Data_Altered$date)$wday %in% c(0,6), 'Week End', 'Week Day')
```

##### 2. Make a panel plot containing a time series plot

```{r}
Average_Activity_Data_Altered <- aggregate(steps ~ interval + dateType, data=Activity_Data_Altered, mean)
ggplot(Average_Activity_Data_Altered, aes(interval, steps)) + 
    geom_line() + 
    facet_grid(dateType ~ .) +
    xlab("Interval") + 
    ylab("Number Of Steps")
```

